---
title: "Reproducible Research"
output: html_document
---
### Peer Assessment I  
Using **knitr** in **R Markdown** to produce document which is answers the assignment questions.  
Assuming "activity.csv" file is in the same directory as the .Rmd file

**Load data and required packages.**

```{r}
suppressMessages(library(ggthemes))
suppressMessages(library(ggplot2))
activity <- read.csv("activity.csv")
activity$date<-as.Date(activity$date)
```

**Project Requirement 1: What is mean total number of steps taken per day?**

```{r}
# Aggregate steps group by date, ignore NA values.
steps_per_day <- aggregate(steps ~ date, data = activity, FUN = sum, na.rm=TRUE)
mean_steps_per_day <- round(mean(steps_per_day$steps))
median_steps_per_day <- round(median(steps_per_day$steps))
```
Mean steps per day = `r as.integer(round(mean(steps_per_day$steps)))`  
Median steps per day = `r as.integer(round(median(steps_per_day$steps)))`  
```{r, fig.width=5, fig.height=3}
ggplot(steps_per_day, aes(x=date,y=steps)) +
    geom_bar(stat = "identity", position=position_dodge(),color="grey") +
    stat_smooth(method="lm", se=FALSE, colour = "red", size = 1) +
    theme_economist() + scale_colour_economist() +
    xlab("Date")  + ylab("Steps per day") 
```

**Project Requirement 2: What is the average daily activity pattern?**

```{r}
activity_cc <- na.omit(activity)
steps_per_interval <- aggregate(steps ~ interval, data=activity_cc,FUN=mean)
max_steps_interval <- round(max(steps_per_interval$steps))
interval_max_steps <- steps_per_interval$interval[which.max(steps_per_interval$steps)]
```
Reporting interval `r interval_max_steps` having maximum activity.  
Average of `r max_steps_interval` steps taken during this interval.  
```{r, fig.width=5, fig.height=3}
ggplot(steps_per_interval, aes(x=interval, y=steps)) +
    geom_line(stat="identity") + theme_economist() + scale_colour_economist() +
    geom_vline(xintercept=steps_per_interval$interval[which.max(steps_per_interval$steps)] ,color="red") +
    labs(title="Activity Pattern",fontsize=1) 

```

**Project Requirement 3: Imputing missing values**  

```{r}
activity2 <- activity
activity2$steps[is.na(activity2$steps)] <- mean(na.omit(activity2$steps))
steps_per_day2 <- aggregate(steps ~ date, data=activity2, FUN=sum)
mean_steps_per_day2 <- round(mean(steps_per_day2$steps))
median_steps_per_day2 <- round(median(steps_per_day2$steps))
```
Mean steps per day after imputing NA's = `r as.integer(round(mean(steps_per_day2$steps)))`  
Median steps per day after imputing NA's = `r as.integer(round(median(steps_per_day2$steps)))`  
```{r, fig.width=5, fig.height=3}
ggplot(steps_per_day2, aes(x=date,y=steps)) +
    geom_bar(stat = "identity", position=position_dodge()) +
    stat_smooth(method="lm", se=FALSE, colour = "red", size = 1) +
    theme_economist() + scale_colour_economist() +
    xlab("Date")  + ylab("Steps per day") +
    labs(title="Steps taken per day - Imputing NA's") 
```
  
**Project Requirement 4: Differences in activity patterns between weekdays and weekends**  

```{r, fig.width=5, fig.height=3}
activity2$day[weekdays(as.Date(activity2$date)) %in% c("Saturday", "Sunday")] <- "Weekend"
activity2$day[!weekdays(as.Date(activity2$date)) %in% c("Saturday", "Sunday")] <- "Weekday"
activity2[, 4] <- as.factor(activity2[, 4])
steps.new <- aggregate(steps ~ interval + day, data=activity2, FUN=mean)
ggplot(steps.new, aes(x=interval, y=steps, group=1)) +
    geom_line() + facet_wrap(~ day, ncol=1) +
    theme_economist() + scale_colour_economist()
```
